---
title: "Lab-D-Replication"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This notebook replicates the Lab-D assignment, which geocodes the locations of chain grocery stores in St. Louis City and County.

## Dependencies
This notebook requires the following packages:

```{r load-packages}
# tidyverse packages
library(dplyr)       # data wrangling
library(readr)       # read csv files
library(stringr)     # data wrangling for strings
library(tidyr)       # data wrangling for tables

# spatial packages
library(censusxy)    # geocoding
library(mapview)     # preview spatial data
library(sf)          # spatial tools

# other packages
library(here)        # file path management
```

## Load Data
This notebook requires raw data on grocery store locations:

```{r load-data}
grocery <- read_csv(here("data", "STL_RETAIL_Grocery.csv"))
```

## Part 1
### Question 1
First, we need to pre-process our data and parse the address colum by:

  1. Parsing the zip code into its own column, **then**
  2. deleting the zip code from the `address` column, **then**
  3. removing trailing spaces, **then**
  4. parsing the remaining address data.

```{r p1-q1}
grocery %>%
  mutate(zip = word(address, -1)) %>% # extract zip code
  mutate(address = str_replace(string = address, pattern = zip, replacement = "")) %>% # delete zip code in address
  mutate(address = str_squish(address)) %>% # remove trailing space
  separate(col = "address", into = c("street", "city", "state"), sep = ",") -> grocery # parse
```

We now have data that can be geocoded!

### Question 2
Now, we'll geocode our address data:

```{r p1-q2}
grocery_geocode <- cxy_geocode(grocery, id = "id", street = "street", city = "city", state = "state", zip = "zip")
```

Notice that we have a lot of `NA` values - this is far less than ideal. Geocoding often takes extensive work to prep and clean our data. 

### Question 3
Finally, we'll project them using NAD 1983:

```{r p1-q3-project}
grocery_geocode %>%
  filter(is.na(cxy_lon) == FALSE) %>%
  st_as_sf(coords = c("cxy_lon", "cxy_lat"), crs = 4269) -> grocery_sf
```

We can preview our data that we could geocode with `mapview`:

```{r p1-q3-explore}
mapview(grocery_sf, zcol = "title")
```
