---
title: "Lecture-D Examples Complete"
author: "Your Name"
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This notebook introduces concepts related to geocoding using `R`.

## Install Additional Dependency
Our geocoding package, `censusxy`, needs to be installed:

```{r install-package, eval = FALSE}
remotes::install_github("slu-openGIS/censusxy")
```

## Dependencies
This notebook requires the follwing packages:

```{r load-packages}
# tidyverse packages
library(dplyr)       # data wrangling
library(readr)       # read csv files
library(stringr)     # data wrangling for strings
library(tidyr)       # data wrangling for tables

# spatial packages
library(censusxy)    # geocoding
library(mapview)     # preview spatial data
library(sf)          # spatial tools

# other packages
library(here)        # file path management
```

## Load Data
This notebook requires two data sets:

```{r load-data}
pharmacy <- read_csv(here("data", "example-data", "STL_RETAIL_Pharmacy.csv"))
pharmacy_simple <- read_csv(here("data", "example-data", "STL_RETAIL_Pharmacy_Simple.csv"))
```

## Preparing Data for Geocoding
Geocoders need data that have been pre-processed into separate columns. Sometimes data come this way, other times they do not. The sample data have a common configuration with the entire street address in a single string. 

### Simple Case
In the simple case, each of the major parts are separated by a comma: street, city, state, and zip. In our `pharmacy_simple` data, our data follow this convention:

```{r preview-simple}
pharmacy_simple
```

We can use the `tidyr` package's `separate()` function to parse these data:

```{r simple-parse}
pharmacy_simple <- separate(pharmacy_simple, col = "address", into = c("street", "city", "state", "zip"), sep = ",")
```

This is useful for illustrating how `separate()` works, and introducing the idea of parsing street addresses. However, it isn't realistic for most US data.

### Realistic Case
In the US, most human-entered data omits the final comma between states and zip-codes. This makes the parsing process slightly more challening. The `pharmacy` data are formatted like this:

```{r preview-realistic}
pharmacy
```

We need to use a couple of tools from `stringr` to help pull the zip code out prior to separating our data. Our goal is to first remove the zip code from our address string and then parse. First, we'll take care of removing the zip code and the final comma.

```{r prep}
pharmacy %>%
  mutate(zip = word(address, -1)) %>% # extract zip code
  mutate(address = str_replace(string = address, pattern = zip, replacement = "")) %>% # delete zip code in address
  mutate(address = str_squish(address)) -> pharmacy # remove trailing space
```

Then we can again use the `tidyr` package's `separate()` function to pull these apart:

```{r parse}
pharmacy <- separate(pharmacy, col = "address", into = c("street", "city", "state"), sep = ",")
```

We now have parsed data!

## Geocoding
Once our data are parsed, we can feed them to `censusxy`. This package accesses the US Census Bureau's geocoding API, which takes American street addresses and attempts to match them with an approximate latitude and longitude. On the plus side, it is free (unlike many other services), and we've built `censusxy` to handle bulk requests (unlike other `R` packages). On the minus side, it is pretty slow.

```{r geocode}
pharmacy_geocode <- cxy_geocode(pharmacy, id = "id", street = "street", city = "city", state = "state", zip = "zip")
```

Notice that we have one `NA` observation - this is common with "real world" data and geocoding to have addresses that cannot be correctly matched.

Next, we'll project them using NAD 1983:

```{r project}
pharmacy_geocode %>%
  filter(is.na(cxy_lon) == FALSE) %>%
  st_as_sf(coords = c("cxy_lon", "cxy_lat"), crs = 4269) -> pharmacy_sf
```

Finally, we can preview our data with `mapview`:

```{r explore}
mapview(pharmacy_sf, zcol = "title")
```



```{r move-to-docs, include=FALSE}
# you do need to include this in any notebook you create for this class
fs::file_copy(here::here("examples", "lecture-D-complete.nb.html"), 
              here::here("docs", "index.nb.html"), 
              overwrite = TRUE)
```
